### 2010-03-30 papers/icfp-2010-desc/paper_idesc.tex
Desc@ICFP: Structural changes to IDesc examples

In this section, we develop more structure for our indexed
data-types. In Section~\ref{sec:desc-free-monad}, we have built a free
monad operation for simple descriptions. Building on this experience,
we are going to present its equivalent in the indexed world. The
process the same. Namely, given an indexed functor, we derive the
indexed functor coding its free monad: \note{pwm: Whoa there. Maybe we
  should say something about IMonads in general before we get to this
  point?}

@@

Just as in the universe of descriptions, this construction comes with
an obvious \return\ and a substitution operation, the \bind. Its type
signature is the following:

@@

In Section~\ref{sec:idesc-examples}, we had the intuition that our
datatypes $\HExprD$ and $\HExprVarD{\V{dom}}$ enjoy a monadic structure. We had
identified the variable substitution operation as the \bind\ of a free
monad. The definition of $\substI$ above confirm our intuition: using
$\substI$, we should easily obtain $\substH$. To do so, we first have
to massage the definition of our data-type, to exhibit its monadic
structure.

As previously mentioned, we identify $\Eval{}$ as the \return\ of the
free monad, while the other components are the action of the monad. As
a result, the definition is similar to $\HExprD$ presented in
Figure~\ref{fig:hexpr-full}, at the exception of $\HExprAD$ that is
replaced by $\HExprAFreeD$:

@@

Hence, we have completed our implementation of the evaluator for open
terms. We started with a well-typed language of arithmetical
expressions. We have seen how to take advantage of indexes to define
type-safe constructors. Then, we have implemented an evaluator for
closed term, based on the generic catamorphism function. Having built
an open term representation, we wanted a substitution operation, in
order to close open terms in a context. Again, we have implemented
this operation with the generic substitution operator. Hence, without
much efforts, we have described the syntax of a well-typed language,
together with its semantics.

>>>>

In this section, we develop more structure for our indexed
data-types. In Section~\ref{sec:desc-free-monad}, we have constructed
a free monad operation on simple descriptions. Building on this
experience, we are going to present its equivalent in the indexed
world. The process the same. Namely, given an indexed functor, we
derive the indexed functor coding its free monad:
\note{pwm: Whoa there. Maybe we should say something about IMonads in
general before we get to this point?}

@@

Just as in the universe of descriptions, this construction comes with
an obvious \return and a substitution operation, the \bind. Its type
signature is the following:

@@

In Section~\ref{sec:idesc-examples}, we had the intuition that our
datatypes $\HExprD$ and $\HExprVarD{\V{dom}}$ enjoy a monadic structure. We had
identified the variable substitution operation as the \bind of a free
monad. The definition of $\substI$ above confirm our intuition: using
$\substI$, we should easily obtain $\substH$. To do so, we first have
to massage the definition of our data-type, to exhibit its monadic
structure.

As previously mentioned, we identify $\Eval{}$ as the \return of the
free monad, while the other components are the action of the monad. As
a result, the definition is similar to $\HExprD$ presented in
Figure~\ref{fig:hexpr-full}, at the exception of $\HExprAD$ that is
replaced by $\HExprAFreeD$:

@@

Hence, we have completed our implementation of the evaluator for open
terms. We started with a well-typed language of arithmetical
expressions. We have seen how to take advantage of indexes to only
present type-safe constructors. Then, we have implemented an evaluator
for closed term, based on the generic catamorphism function. Having
implemented an open term representation, we wanted a substitution
operation, in order to close open terms in a context. Again, we have
implemented this operation with the generic substitution
operator. Hence, without much efforts, we have described the syntax of
a well-typed language, together with its semantics.
